\section{Preliminaries}
Here a couple of preliminariy definitions, before we move on to categories.

\begin{defn}[$R$-module]
	A {\em left} $R$-module $\left(M, +, \cdot\right)$ is an abelian group $\left(M, +\right)$ on which is defined a map
	\begin{align}
		\cdot\colon R \cross M &\to M \\
		\left(r, m\right) &\mapsto rm
	.\end{align} 
	It is called scalar multiplication and satisfies
	\begin{enumerate}
		\item for any $r \in R$ the induced map
			 \begin{align}
				\dot{r}\colon M &\to M \\
				m &\mapsto rm
			\end{align} 
			is a homomorphism of abelian groups.
		\item The map that sends each $r \in R$ to its associated endomorphism (as in the above)
			\begin{align}
				\phi\colon R &\to \mathrm{End}_{\Z}\left( M \right) \\
				r &\mapsto \dot{r}
			\end{align} 
			is a morphism of rings.
	\end{enumerate}
	If $\phi$, instead of being a homomorphism is an antihomomorphism (i.e. it is a homomorphism
	\begin{align}
		\phi\colon R^{op} &\to \mathrm{End}_{ \Z}\left( M \right)
	,\end{align} 
	from the opposite ring, in which the operations are computed in the opposite direction), then $M$ is a {\em right} $R$-module.
	We denote {\em left} $R$-modules as $\prescript{}{R}{M}$, wheareas {\em right} $R$-modules as $M_R$.
\end{defn}

\begin{defn}[Bimodule]
	Let $R,S$ be rings.
	An abelian group $\left(M, +\right)$ is an $R,S$-bimodule ${}_{R}{M}_{S}$ iff 
	$\prescript{}{R}{M}$ is a left $R$-module, $M_S$ is a right $R$-module and
	\begin{equation}
		r(xs) = (rx)s
	\end{equation} 
	for any $r \in R$, $s \in S$, $x \in M$.
\end{defn}

\begin{defn}[Tensor product of modules]
	Let $S$ be a ring, $M_S \in \mathsf{Mod}\text{-}S$ and ${}_SN \in S\text{-}\mathsf{Mod}$.
	A map $\beta\colon M \cross N \to G$, to $G$ an abelian group, is called balanced iff
	it satisfies the following
	\begin{align}
		\beta(m + m', n) &= \beta(m,n) + \beta(m',n) \qquad & \,\forall\, m, m' \in M \text{ and } \,\forall\, n \in N\\
		\beta(m, n + n') &= \beta(m,n) + \beta(m,n') \qquad & \,\forall\, m \in M \text{ and } \,\forall\, n, n' \in N\\
		\beta(ms, n) &= \beta(m,sn) \qquad & \,\forall\, m \in M,\ \,\forall\, n \in N \text{ and } \,\forall\, s \in S
	.\end{align} 
	The tensor product of $M$ and $N$ is the pair $\left(M \otimes_{S} N, \tau \right)$,
	with $M \otimes_{S} N$ an abelian group and $\tau\colon M \cross N \to M \otimes_{S} N$ a map s.t.
	$\,\forall\, \beta\colon M \cross N \to G$ a balanced map, $\exists\, !\, \alpha\colon M \otimes_{S} N \to G$ an abelian group morphism
	s.t. the following diagram commutes
	\begin{equation}
	\begin{tikzcd}
		M \cross N \arrow[r, "\tau", rightarrow] \arrow[d, "\beta"', rightarrow] &
		M \otimes_{S} N \arrow[dl, "\alpha", rightarrow] \\
		G &
	\end{tikzcd}
	,\end{equation} 
	i.e. s.t. $\alpha \circ \tau = \beta$.
	In such a case we say that every balanced map $\beta$ factors through $\tau$ via an abelian group morphism.
\end{defn}

\begin{rem}[Construction of the tensor product]
	Consider $M$ and $N$ as before.
	Consider $M \cross N$ as a set.
	Let $\Z^{M \cross N}$ be the free abelian group with basis $\left(m, n\right) \in M \cross N$.
	Consider $H \triangleleft \Z^{M \cross N}$ generated by the elements of the form
	\begin{align}
		\big\{& \left(m + m', n\right) - \left(m, n\right) - \left(m',n \right),
	\left(m, n + n'\right) - \left(m, n\right) - \left(m, n'\right),\\
		      &\left(ms, n\right) - \left(m, sn\right)
		     \ \big|\ m,m' \in M,\ n,n' \in N,\ s \in S \big\}
	.\end{align} 
	Let $\tau\colon M \cross N \to \Z^{M \cross N}/H$, defined by $\left(m, n\right) \xmapsto{\tau} \left(m, n\right) + H$,
	then $\left(\Z^{M \cross N}, \tau\right)$ is a tensor product of $M$ and $N$.
\end{rem}

\begin{rem}[Tensor product as a module]
	Given $N_S \in \mathsf{Mod}\text{-}S$ and ${}_S M_R$ and $S$-$R$ bimodule,
	we want to construct a tensor product of the two, which is also a right $R$-module.
	(We can choose $S = \mathrm{End}_{R}\left( M \right)$, or $S = \Z$ for example).
	We define, for any $L_R \in \mathsf{Mod}\text{-}R$, the set
	\begin{equation}
		\beta \in \mathrm{Bal}(N \cross M, L_R)
	,\end{equation} 
	consisting of $\beta$ balanced maps (as defined above) s.t. $\beta(x, yr) = \beta(x,y) r$
	for all $x \in N$, $y \in M$ and $r \in R$.
Then, in this situation, we define the tensor product $N \otimes_{S} M$ as the right $R$-module, with a map $\tau$, s.t.
the following diagram commutes
\begin{equation}
\begin{tikzcd}
	N \cross M \arrow[r, "\tau", rightarrow] \arrow[d, "\beta"', rightarrow] &
	N \otimes_S M \arrow[dl, "\exists\, \alpha", dashrightarrow] \\
	L_R &
\end{tikzcd}
\end{equation} 
for all $L_R$, and $\beta \in \mathrm{Bal}(N \cross M, L_R)$.
In particular this gives a bijection
\begin{equation}
\begin{tikzcd}
	\mathrm{Bal}(N \cross M, L_R) \arrow[r, "\varphi", leftrightarrow] &
	\mathrm{Hom}_{R} \left( N \otimes_S M, L_R \right)
\end{tikzcd}
.\end{equation} 
\end{rem}

